Lyman-alpha spectra from rotating outflows

### 2018-03-27 paper/paper.tex
referencias
edit

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these kinematic
features impact the morphology of the Lyman-$\alpha$ emission line.
We find three important consequences of rotation.
First, it increases the detected flux at the line's center in comparison with the
non-rotational case, second it produces a line broadening and third it introduces
a dependency with the  viewing angle.
We also develp a semi-analytic model that modifies the spectra of
pure outflow simulations that manages to reproduce the main features of the
Monte Carlo simulations.
As an application we constraint the rotation parameters of a set of observed LAE
emission spectra.
\end{abstract}

@@

Dwarf galaxies show signs of coherent rotation
\citep{2009A&A...493..871S}.
Some of them are expected to have high neutral gas contents and thus
show \lya\ emission with rotation imprints
\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
Some Compact Dwarf Galaxies actually show gas kinematics that range
from pure rotation to high velocity dispersion without a clear
rotation pattern.
\citep{2015A&A...577A..21C,2017arXiv170809407C,2017A&A...600A.125C}

@@

CLARA follows the propagation of individual photons through a neutral
Hydrogen medium characterized by its temperature, velocity field and
global optical depth.
The code assumes an homogeneous density throughout the simulated
volume.
In the current implementation we neglect the influence of dust.
Our basic model is an spherical distribution of neutral hydrogen,
an approximation commonly used in the literature, as it explains a
wide variety of observational features \citep{Ahn03,Verhamme06,Dijkstra06}.

The central element in this paper is the velocity field that captures
outflows and rotation.
Outflows are captured by a Hubble-like radial velocity profile with
the velocity magnitude increasing linearly with the radial
coordinate; the outflows model is fully characterized by $V_{\rm out}$, the
velocity at the sphere's surface.
Rotation follows a solid body rotation profile, which is fully
characterized by $V_{\rm rot}$, the linear velocity at the sphere's surface.

@@

\subsection{Analytical Model for Bulk Rotation}

Rotation induces two main effects on the \lya line morphology.
First. Break of spherical symmetry due to the
preferential direction introduced by the rotation axis.
This symmetry break is reflected in the observed spectra as a
dependency on the viewing angle, $\theta$, between the line of sight
of a distant observer and the rotational velocity.
Second. Line blurring as the rotational velocity increases.
At a fixed viewing angle, faster rotation induces makes the observed
line wider and increases the flux around the line's center.

\cite{Garavito14} presented in the Appendix an analytical model that
accounts for these two features.
The basic assumption of the analytical model is that each
differential surface element on the sphere Doppler shifts the photons
that it emits
In this paper we introduce this ansatz by post-processing the results
of the outflows simulations without rotation.
This allows us to produce new \lya spectra and compare them with the
full radiative transfer solution including both outflows and
rotation.

@@

Rotation alone does have an impact on the \lya line morphology
and we obtain the same influence as in the literature.
Together with the outflows, we found that the resulting spectra is
equivalent to a non-rotating galaxy with outflows that undergoes a
Doppler Shift in velocity, weighted by the amount of photons that if
rotating would project their escape in $\hat{k}$'s direction:

\begin{equation}
V' = V + \vec{v}_{\rm rot}\cdot\hat{k}
\label{eq:shift_x}
\end{equation}
%
where $V'$ is the new Doppler shifted velocity, $V$ is the velocity caused by
outflows.

>>>>

\begin{abstract}
Outflows and rotation are two ubiquitous kinematic features in the gas
kinematics of galaxies.
Here we perform Monte Carlo radiative transfer simulations of outflowing
gas with additional solid body rotation to understand how these kinematic
features impact the morphology of the Lyman-$\alpha$ emission line.
We find three important consequences of rotation.
First, it increases the detected flux at the line's center in comparison with the
non-rotational case, second it produces a line broadening and third it introduces
a dependency with the  viewing angle.
We also develp a semi-analytic model that modifies the spectra of
pure outflow simulations that manages to reproduce the main features of the
Monte Carlo simulations.
As an application to observational data we constraint the rotation
parameters of a set of observed LAE emission spectra.    
\end{abstract}

@@

For instances, dwarf galaxies show signs of coherent rotation
\citep{2009A&A...493..871S}.
Some of them are expected to have high neutral gas contents and thus
show \lya\ emission with rotation imprints
\citep{2005A&A...433L...1B,2008ApJ...672..888T,2013MNRAS.434.2491G}
Some Compact Dwarf Galaxies actually show gas kinematics that range
from pure rotation to high velocity dispersion without a clear
rotation pattern.
\citep{2015A&A...577A..21C,2017A&A...600A.125C}

@@

The Monte Carlo code we use (CLARA)  follows the propagation of
individual photons through a neutral Hydrogen medium characterized by
its temperature, velocity field and global optical depth.
The code assumes an homogeneous density throughout the simulated
volume.
In the current implementation we neglect the influence of dust.
Our basic model is an spherical distribution of neutral hydrogen,
an approximation commonly used in the literature, as it explains a
wide variety of observational features
\citep{Ahn03,Verhamme06,Dijkstra06}.

The velocity field we use captures both outflows and rotation.
Outflows are described by a Hubble-like radial velocity profile with
the velocity magnitude increasing linearly with the radial
coordinate; the outflows model is fully characterized by $v_{\rm out}$, the
velocity at the sphere's surface.
Rotation follows a solid body rotation profile, which is fully
characterized by $v_{\rm rot}$, the linear velocity at the sphere's surface.

@@

\cite{Garavito14} presented in the Appendix an analytical model that
accounts for the effects of pure rotation on the
\lya\ lines 'morphology. 
The basic assumption of their analytical model is that each
differential surface element on the sphere Doppler shifts the photons
that it emits

In this paper we introduce this ansatz by post-processing the results
of the outflows simulations without rotation.
We introduce a Doppler shift as follows
\begin{equation}
x' = x + \frac{\vec{v}_{\rm rot}\cdot\hat{k}}{v_{\rm th}}
\label{eq:shift_x}
\end{equation}
%
where $x'$ is the photon's new frequency, $x$ is the photon's
frequency after being procesed by the outflow, $v_{\rm rot}$ is the 
rotational velocity at the point of escape of the photon, $\hat{k}$ is
the photon's direction of propagation and $v_{\rm th}$ is the thermal
velocity of the sphere.

This allows us to produce new \lya spectra and compare them with the
full radiative transfer solution including both outflows and
rotation.

@@
